---
title: "Wildfires"
subtitle: "Simulated Dataset"
date: "`r format(Sys.time(), '%d %B, %Y')`"

format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    link-external-newwindow: true

execute:
  warning: false
  
from: markdown+emoji  
---

```{r}
#| label: setup
#| echo: false

# load package(s)
library(tidyverse)

# load datasets
wf <- read_csv("wildfires.csv")

ranger <- tibble(
  x = c(10, -5, -5),
  y = c(5, 2, -8)
)

wildlife <- tibble(
  x = c(15, 18, 18, 16.5, 15),
  y = c(16, 16, 18, 19, 18)
)
```

## Description

This data describes 500 wildfires that started within a large national park. The origin of each fire is shown below in red.

```{r}
#| label: plot
#| echo: false


ggplot(wf) +
  geom_point(
    mapping = aes(x, y),
    color = "firebrick2",
    shape = 8
  ) +
  geom_point(
    data = ranger,
    mapping = aes(x, y),
    color = "forestgreen",
    size = 4,
    shape = 17
  ) +
  geom_polygon(
    data = wildlife,
    mapping = aes(x, y),
    fill = "cyan3"
  ) +
  labs(x = NULL, y = NULL) +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "navajowhite2")
  ) +
  annotate(
    "text",
    x = 17.25, y = -15,
    label = "Ranger Station",
    hjust = 1
  ) +
  annotate(
    "point",
    x = 18.25, y = -15,
    shape = 17,
    color = "forestgreen",
    size = 5
  ) +
  annotate(
    "text", 
    x = 17.25, y = -13.1, 
    label = "Wildlife Protection Zone", 
    hjust = 1
    ) +
  annotate(
    "point", 
    x = 18.25, y = -13.1, 
    shape = 16, 
    color = "cyan3", 
    size = 5
    ) +
  annotate(
    "text", 
    x = 17.25, y = -11.2, 
    label = "Fire Origin", 
    hjust = 1
    ) +
  annotate(
    "point", 
    x = 18.25, y = -11.2, 
    shape = 8, 
    color = "firebrick2", 
    size = 5
    )
```

## Data

The dataset contains observations on 500 fires started in a large national park. For each fire, the data contain 15 variables, in addition there are two outcomes ('burned' and 'wlf'). These variables are:

-   **`r names(wf)[1]`**, and **`r names(wf)[2]`**: the (x, y) coordinates of where the fire started.
-   **`r names(wf)[3]`**: air temperature when the fire started
-   **`r names(wf)[4]`**: air humidity when the fire started
-   **`r names(wf)[5]`**: wind speed when the fire started
-   **`r names(wf)[6]`**: wind direction when the fire started
-   **`r names(wf)[7]`**: rainfall in the week preceding the fire
-   **`r names(wf)[8]`**: number of days since the last fire
-   **`r names(wf)[9]`**: amount of vulnerable (unburned) foliage near the fire
-   **`r names(wf)[10]`**: indicates if another fire is ongoing when the fire starts
-   **`r names(wf)[11]`**: indicates if the fire started near a ranger station, and that ranger station was manned. Note that even if a fire started near a ranger station, that station may not have been manned at the time.
-   **`r names(wf)[12]`**: indicates if the fire occurred before 1950, which it became more common to use planes and helicopters to fight forest fires.
-   **`r names(wf)[13]`**: indicates if a helicopter was available to fight the fire
-   **`r names(wf)[14]`**: composite score for the resources available to fight the fire, including manpower and available air support
-   **`r names(wf)[15]`**: indicates the level of foot traffic in the national park when the fire started
-   **`r names(wf)[16]`**: number of hectares burned by the fire
-   **`r names(wf)[17]`**: indicates fire reached the wildlife protection zone
