---
title: "Group Project 1"
author: "STAT 301-1"
date: ''
output:
  html_document:
    df_print: paged
---

```{r echo=F, message=FALSE, warning=FALSE}
library(tidyverse)
wf <- read_csv("wildfires.csv")
ranger <- tibble(x = c(10, -5, -5), y = c(5, 2, -8))
wildlife <- tibble(x = c(15, 18, 18, 16.5, 15), y = c(16, 16, 18, 19, 18))
```

## Desciption

For the group project, you will be using the wildfires dataset (wildfires.csv). This data describes 500 wildfires that started within a large national park. The origin of each fire is shown below in red. 

```{r, echo=F}
ggplot(wf) +
  geom_point(aes(x, y), color = "firebrick2", shape = 8) +
  geom_point(data = ranger, aes(x, y), color = "forestgreen", size = 4, shape = 17) +
  geom_polygon(data = wildlife, aes(x, y), fill = "cyan3") +
  labs(x = "", y = "") +
  theme(
    axis.ticks = element_blank(), 
    axis.text = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_rect(fill = "navajowhite2")
    ) +
  annotate("text", x = 17.25, y = -15, label = "Ranger Station", hjust = 1) +
  annotate("point", x = 18.25, y = -15, shape = 17, color = "forestgreen", size = 5) +
  annotate("text", x = 17.25, y = -13.1, label = "Wildlife Protection Zone", hjust = 1) +
  annotate("point", x = 18.25, y = -13.1, shape = 16, color = "cyan3", size = 5) +
  annotate("text", x = 17.25, y = -11.2, label = "Fire Origin", hjust = 1) +
  annotate("point", x = 18.25, y = -11.2, shape = 8, color = "firebrick2", size = 5)
```

## Data

The dataset contains observations on 500 fires started in a large national park. For each fire, the data contain 15 variables, in addition there are two outcomes ('burned' and 'wlf'). These variables are:

  * __`r names(wf)[1]`__, and __`r names(wf)[2]`__: the (x, y) coordinates of where the fire started.
  * __`r names(wf)[3]`__: air temperature when the fire started
  * __`r names(wf)[4]`__: air humidity when the fire started
  * __`r names(wf)[5]`__: wind speed when the fire started
  * __`r names(wf)[6]`__: wind direction when the fire started
  * __`r names(wf)[7]`__: rainfall in the week preceding the fire
  * __`r names(wf)[8]`__: number of days since the last fire
  * __`r names(wf)[9]`__: amount of vulnerable (unburned) foliage near the fire
  * __`r names(wf)[10]`__: indicates if another fire is ongoing when the fire starts
  * __`r names(wf)[11]`__: indicates if the fire started near a ranger station, and that ranger station was manned. Note that even if a fire started near a ranger station, that station may not have been manned at the time.
  * __`r names(wf)[12]`__: indicates if the fire occurred before 1950, which it became more common to use planes and helicopters to fight forest fires.
  * __`r names(wf)[13]`__: indicates if a helicopter was available to fight the fire
  * __`r names(wf)[14]`__: composite score for the resources available to fight the fire, including manpower and available air support
  * __`r names(wf)[15]`__: indicates the level of foot traffic in the national park when the fire started
  * __`r names(wf)[16]`__: number of hectares burned by the fire
  * __`r names(wf)[17]`__: indicates fire reached the wildlife protection zone
    
